Supplement: Analysis and code for 'Testing Perceptual Accuracy in Charts using Surveys

### 2023-07-20 analysis/analysis.qmd
floating bars

@@

@@

>>>>

```{r data-rnd5}
# other github repo. 
round5 <- readRDS(file.path(vs_repo_local, "round5.rds"))

# floating bars (unaligned) are shown to all participants of round 5
round5_adj <- round5 %>% mutate(
  ab = ab1,
  ab_zoom = ab1_zoom,
  ab_certain = ab1_certain,
  weight = WEIGHT_NOLA
) %>% filter(
  ab != "SKIPPED ON WEB"
)
```

@@

## Model 4: Unaligned Elements in horizontal wide barcharts compared to floating bars

```{r}
#| label: fig-context-bars
#| fig-cap: Floating bars show and horizontal wide barcharts show the same areas  with the same distances
#| fig-subcap: 
#|   - floating bars
#|   - horizontal wide
#| layout-nrow: 1
#| out-width: 100%
knitr::include_graphics("images/floating - all.PNG")
knitr::include_graphics("images/round3/hwide-ab.PNG")
```

```{r combine-r35}
# wide horizontal
wide_horiz <- round3_adj %>% filter(design=="Horizontal wide")

# floating bars are in round5
lambda <- calculate_lambda(wide_horiz$weight, round5_adj$weight)

round_35 <- rbind(
  wide_horiz %>% mutate(
    weight = lambda*weight/sum(weight),
    task = "horizontal wide"
  ) %>% dplyr::select(CaseId, weight, ab, ab_zoom, ab_certain, task),
  round5_adj %>% mutate(
    weight = (1-lambda)*weight/sum(weight),
    task = "floating"
  ) %>% dplyr::select(CaseId, weight, ab, ab_zoom, ab_certain, task)
) %>% mutate(
  correct = ab=="B is bigger",
  not_wrong = ab != "A is bigger",
  task = factor(task)
)

survey_35 <- svydesign(ids=~CaseId, weights=~weight, data = round_35)
```

```{r model-r35}
logit_35_correct <- svyglm(correct~task-1, design = survey_35, family=quasibinomial())
logit_35_not_wrong <- svyglm(not_wrong~task-1, design = survey_35, family=quasibinomial())

coefs_35_correct <- broom::tidy(logit_35_correct) %>% 
  mutate(task = factor(gsub("task", "", term), levels=levels(round_35$task)))
coefs_35_notwrong <- broom::tidy(logit_35_not_wrong) %>% 
  mutate(task = factor(gsub("task", "", term), levels=levels(round_35$task)))
```

```{r}
#| label: fig-round-floats
#| fig-cap: "Accuracy of responses by design. Free floating bars are comparisons of lengths. The additional context of stacked barcharts anchors the bars for a significant  increase in correct responses. However, the percentage of wrong responses also significantly increases. "
#| out-width: 70%

round_35 %>% 
  mutate(
    response3 = c("Wrong", "They are the same", "Correct")[correct+not_wrong+1],
    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong")))
  ) %>% 
  ggplot(aes(x = task)) + 
  geom_bar(aes(fill = response3, weight = weight), position="fill") +
#  facet_grid(.~task) + 
  theme(legend.position="bottom") + 
  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
  xlab("") +
  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25)) + 
  geom_rect(
    aes(
      xmin = as.numeric(task) - 0.45,
      ymin = inv_logit(estimate - 1.96* std.error),
      xmax = as.numeric(task) + 0.45,
      ymax = inv_logit(estimate + 1.96* std.error),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = coefs_35_correct
  ) +
  geom_segment(
    aes(
      x = as.numeric(task) - 0.45,
      y = inv_logit(estimate ),
      xend = as.numeric(task) + 0.45,
      yend = inv_logit(estimate )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = coefs_35_correct
  ) +
  geom_rect(
    aes(
      xmin = as.numeric(task) - 0.45,
      ymin = inv_logit(estimate - 1.96* std.error),
      xmax = as.numeric(task) + 0.45,
      ymax = inv_logit(estimate + 1.96* std.error),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = coefs_35_notwrong
  ) +
  geom_segment(
    aes(
      x = as.numeric(task) - 0.45,
      y = inv_logit(estimate ),
      xend = as.numeric(task) + 0.45,
      yend = inv_logit(estimate )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = coefs_35_notwrong
  )
#+ 
#  geom_text(aes(label=letters), y = -0.05, data = ci_confint, colour = "steelblue") +
#  geom_text(aes(label=letters), y = 1.05, data = ci_confint3, colour = "darkorange")

@@

How do our findings relate to @cejaTruthSquareAspect2021 ?

Science communication review of visuals: @desnoyersTaxonomyVisualsScience2023
